\documentclass[12pt]{article}
\usepackage{pmmeta}
\pmcanonicalname{OrderofARing}
\pmcreated{2013-03-22 17:10:31}
\pmmodified{2013-03-22 17:10:31}
\pmowner{Wkbj79}{1863}
\pmmodifier{Wkbj79}{1863}
\pmtitle{order (of a ring)}
\pmrecord{15}{39488}
\pmprivacy{1}
\pmauthor{Wkbj79}{1863}
\pmtype{Definition}
\pmcomment{trigger rebuild}
\pmclassification{msc}{16-01}
\pmsynonym{order}{OrderofARing}
\pmsynonym{order of a ring}{OrderofARing}
\pmrelated{OrderGroup}
\pmdefines{finite ring}

\endmetadata

\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{pstricks}
\usepackage{psfrag}
\usepackage{graphicx}
\usepackage{amsthm}
%%\usepackage{xypic}

\begin{document}
The \emph{order} of a ring $R$ is the \PMlinkname{order}{OrderGroup} of its additive group, \PMlinkname{i.e.}{Ie} the number of elements of $R$.  The order of $R$ can be denoted as $\vert R \vert$.  If $\vert R \vert$ is finite, then $R$ is said to be a \emph{finite ring}.

This definition of order is not necessarily standard.  Please see \PMlinkexternal{this correction}{http://planetmath.org/?op=getobj&from=corrections&id=12149} and the posts attached to it for more details.

This definition of order is used in the following works:

\begin{enumerate}
\item Angerer, Josef and Pilz, G\"{u}nter.  ``The Structure of Near Rings of Small Order.''  \emph{Computer Algebra: EUROCAM '82, European Computer Algebra Conference; Marseilles, France, April 1982}.  Editors: Goos, G. and Hartmanis, J.  Berlin: Springer-Verlag, 1982, pp. 57-64.

\item Buck, Warren.  \emph{\PMlinkexternal{Cyclic Rings}{http://planetmath.org/?op=getobj&from=papers&id=336}}.  Charleston, IL: Eastern Illinois University, 2004.

\item Fine, Benjamin.  ``Classification of Finite Rings of Order $p^2$.''  \emph{Mathematics Magazine}, vol. 66 \#4.  Washington, DC: Mathematical Association of America, 1993, pp. 248-252.

\item Fletcher, Colin R.  ``Rings of Small Order.''  \emph{The Mathematical Gazette}, vol. 64 \#427.  Leicester, England: The Mathematical Association, 1980, pp. 9-22.

\item Lam, Tsi-Yuen.  \emph{A First Course in Noncommutative Rings}.  New York: Springer-Verlag, 2001.

\item Mitchell, James.  \emph{School of Mathematics and Statistics: MT4517 Rings and Fields, Lecture Notes 1}.  St. Andrews, Scotland: University of St. Andrews, 2006.  URL: \PMlinkexternal{http://www-history.mcs.st-and.ac.uk/~jamesm/teaching/MT4517/MT4517-notes1.pdf}{http://www-history.mcs.st-and.ac.uk/~jamesm/teaching/MT4517/MT4517-notes1.pdf}

\item N\"{o}bauer, Christof.  \emph{Numbers of rings on groups of prime power order}.  Linz, Austria: Johannes Kepler Universit\"{a}t Linz.  URL: \PMlinkexternal{http://www.algebra.uni-linz.ac.at/~noebsi/ringtable.html}{http://www.algebra.uni-linz.ac.at/~noebsi/ringtable.html}

\item Schwabe, Eric J. and Sutherland, Ian M.  ``Efficient Mappings for Parity-Declustered Data Layouts.''  \emph{Computing and Combinatorics: 9th Annual International Conference, COCOON 2003; Big Sky, MT, USA, July 2003; Proceedings}.  Editors: Warnow, Tandy and Zhu, Binhai.  Berlin: Springer-Verlag, 2003, pp. 252-261.
\end{enumerate}
%%%%%
%%%%%
\end{document}
